A Tidy Framework and Infrastructure to Systematically Assemble Spatio-temporal Indexes from Multivariate Data

### 2023-07-24 tidyindex.qmd
vimeo link to idx tour video

```{r fig-idx-tour, fig.align='center', fig.height= 12, fig.width=12}
#| fig-cap: "Five frames selected from varying the linear weights of four dimensions in Global Gender Gap Index. The weights vary slightly from the official simple average weights (0.25, 0.25, 0.25, 0.25) to observe how the index and ranking reponse. Full animation is available at [vimeo link]."
frames <- c("001", "013", "037", "061", "085")
ani <- paste0(here::here("figures/"), "idx-tour/", "idx-tour-", frames, ".png")
rl <- lapply(ani, png::readPNG)
gl <-  lapply(rl, grid::rasterGrob)
wrap_plots(gl, ncol = 3) 
```

>>>>

```{r fig-idx-tour, fig.align='center', fig.height= 12, fig.width=12}
#| fig-cap: "Five frames selected from varying the linear weights of four dimensions in Global Gender Gap Index. The weights vary slightly from the official simple average weights (0.25, 0.25, 0.25, 0.25) to observe how the index and ranking reponse. Full animation is available at https://vimeo.com/847874016?share=copy."
frames <- c("001", "013", "037", "061", "085")
ani <- paste0(here::here("figures/"), "idx-tour/", "idx-tour-", frames, ".png")
rl <- lapply(ani, png::readPNG)
gl <-  lapply(rl, grid::rasterGrob)
wrap_plots(gl, ncol = 3) 
```

### 2023-07-24 tidyindex.tex
vimeo link to idx tour video

\thanks{CONTACT: H. Sherry
Zhang. Email: \href{mailto:huize.zhang@monash.edu}{\nolinkurl{huize.zhang@monash.edu}}. }
\begin{document}
\captionsetup{labelsep=space}
\maketitle
\textsuperscript{1} Department of Econometrics and Business
Statistics, Monash University, Melbourne,
Victoria, Australia\\ \textsuperscript{2} Institute of
Statistics, University of Natural Resources and Life
Sciences, Vienna, Austria\\ \textsuperscript{3} Department of
Mathematical Sciences, BNU-HKBU United International College, Zhuhai,
Guangdong, China
\begin{abstract}
\begin{itemize}
\tightlist
\item
  indexes, useful, quantify severity, early monitoring,
\item
  A huge number of indexes have been proposed by domain experts,
  however, a large majority of them are not being adopted, reused, and
  compared in research or in practice.
\item
  One of the reasons for this is the plenty of indexes are quite complex
  and there is no obvious easy-to-use implementation to apply them to
  user's data.
\item
  The paper describes a general pipeline framework to construct indexes
  from spatio-temporal data,
\item
  This allows all the indexes to be constructed through a uniform data
  pipeline and different indexes to vary on the details of each step in
  the data pipeline and their orders.
\item
  The pipeline proposed aim to smooth the workflow of index construction
  through breaking down the complicated steps proposed by various
  indexes into small building blocks shared by most of the indexes.
\item
  The framework will be demonstrated with drought indexes as examples,
  but appliable in general to environmental indexes constructed from
  multivariate spatio-temporal data
\end{itemize}
\end{abstract}
\begin{keywords}
\def\sep{;\ }
indexes\sep data pipeline\sep 
software design
\end{keywords}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, breakable, enhanced, boxrule=0pt, frame hidden]}{\end{tcolorbox}}\fi

@@

Index analysts could be interested in understanding how the index values
and country ranking change when the weights assigned the each dimension
is slightly changed. This can help assess the sensitivity of the country
ranking, test the robustness of the indexes to different variables, and
identify the strengths and weaknesses of countries in each aspect. We
run a local tour focusing on the four dimensions of countries in the
South Asia and Sub-Saharan Africa region. The local tour generates an
animation starting with an equal weight of 0.25 for all four dimensions
and gradually increases and reduces the weight of one variable at a
time. In Figure~\ref{fig-idx-tour}, we select five frames from the tour
animation. The first frame represents equal weights, and the subsequent
four frames have a higher weight for each of the four dimensions. When
health receives a higher weight, the impact on the index value is
minimal. Index values and ranking change slightly when the dimension
economy and education are given a greater weight, highlighting countries
with a relatively better (moving right) or worse (moving left) in these
single dimensions. For example, when economy is given a higher weight,
the index moves to the left for Ethiopia and Senegal, revealing it as a
weakness for these two countries to improve compared to their peers with
similar rankings. These information will not be observed when using a
single equal weight scheme since the advantageous and disadvantageous
dimension of a country will be averaged out.

@@

\caption{\label{fig-idx-tour}Five frames selected from varying the
linear weights of four dimensions in Global Gender Gap Index. The
weights vary slightly from the official simple average weights (0.25,
0.25, 0.25, 0.25) to observe how the index and ranking reponse. Full
animation is available at {[}vimeo link{]}.}

>>>>

\thanks{CONTACT: H. Sherry
Zhang. Email: \href{mailto:huize.zhang@monash.edu}{\nolinkurl{huize.zhang@monash.edu}}. }
\begin{document}
\captionsetup{labelsep=space}
\maketitle
\textsuperscript{1} Department of Econometrics and Business
Statistics, Monash University, Melbourne,
Victoria, Australia\\ \textsuperscript{2} Institute of
Statistics, University of Natural Resources and Life
Sciences, Vienna, Austria\\ \textsuperscript{3} Department of
Mathematical Sciences, BNU-HKBU United International College, Zhuhai,
Guangdong, China
\begin{abstract}
\begin{itemize}
\tightlist
\item
  indexes, useful, quantify severity, early monitoring,
\item
  A huge number of indexes have been proposed by domain experts,
  however, a large majority of them are not being adopted, reused, and
  compared in research or in practice.
\item
  One of the reasons for this is the plenty of indexes are quite complex
  and there is no obvious easy-to-use implementation to apply them to
  user's data.
\item
  The paper describes a general pipeline framework to construct indexes
  from spatio-temporal data,
\item
  This allows all the indexes to be constructed through a uniform data
  pipeline and different indexes to vary on the details of each step in
  the data pipeline and their orders.
\item
  The pipeline proposed aim to smooth the workflow of index construction
  through breaking down the complicated steps proposed by various
  indexes into small building blocks shared by most of the indexes.
\item
  The framework will be demonstrated with drought indexes as examples,
  but appliable in general to environmental indexes constructed from
  multivariate spatio-temporal data
\end{itemize}
\end{abstract}
\begin{keywords}
\def\sep{;\ }
indexes\sep data pipeline\sep 
software design
\end{keywords}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, enhanced, breakable, interior hidden, sharp corners]}{\end{tcolorbox}}\fi

@@

Index analysts could be interested in understanding how the index values
and country ranking change when the weights assigned the each dimension
is slightly changed. This can help assess the sensitivity of the country
ranking, test the robustness of the indexes to different variables, and
identify the strengths and weaknesses of countries in each aspect. We
run a local tour focusing on the four dimensions of countries in the
South Asia and Sub-Saharan Africa region. The local tour generates an
animation starting with an equal weight of 0.25 for all four dimensions
and gradually increases and reduces the weight of one variable at a
time. In Figure~\ref{fig-idx-tour}, we select five frames from the tour
animation: the first frame represents equal weights, and the subsequent
four frames have a higher weight for each of the four dimensions. When
health receives a higher weight, the impact on the index value is
minimal. Index values and ranking change slightly when the dimension
economy and education are given a greater weight, highlighting countries
with a relatively better (moving right) or worse (moving left) in these
single dimensions. For example, when economy is given a higher weight,
the index moves to the left for Ethiopia and Senegal, revealing it as a
weakness for these two countries to improve compared to their peers with
similar rankings. These information will not be observed when using a
single equal weight scheme since the advantageous and disadvantageous
dimension of a country will be averaged out.

@@

\caption{\label{fig-idx-tour}Five frames selected from varying the
linear weights of four dimensions in Global Gender Gap Index. The
weights vary slightly from the official simple average weights (0.25,
0.25, 0.25, 0.25) to observe how the index and ranking reponse. Full
animation is available at https://vimeo.com/847874016?share=copy.}
